---
title: "Energy Decomposition Analysis Using `EDA` R Package"
author: "Yongze Song"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Energy Decomposition Analysis Using `EDA` R Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Energy decomposition analysis is used for measuring contributions of components and factors on energy or carbon emission changes ([Ang 2005][2], [Marlay 1984][3], [Liu et al. 2003][4]). 

**EDA** package consists of five energy consumption analysis methods: Log Mean Devisia Index (LMDI) method ([Ang 2005][2]), Laspeyres method, Paasche method (Paasche 1874), Marshall-Edgeworth method (Marshall 1887, Edgeworth 1925) and Walsh method (Walsh 1921). The formulas of five methods are listed in Table 1. 

Table 1. 

Method | use in `EDA` function | Formula
--- | ----------------------- | ---
LMDI | `method = "LMDI"` | \({\Delta} C_{x_{1}} = \sum_{i} L\left(C_{i}^{T}, C_{i}^{0}\right) \ln\frac{x_{i}^{T}}{x_{i}^{0}} = \sum_{i}\frac{C_{i}^{T}-C_{i}^{0}}{\ln C_{i}^{T} - \ln C_{i}^{0}} \ln\frac{x_{i}^{T}}{x_{i}^{0}} \)
Laspeyres | `method = "Laspeyres"` | \({\Delta} C_{x_{1}} = \sum_{i} \left(x_{1i}^{T} - x_{1i}^{0}\right) x_{2i}^{0} ... x_{ni}^{0}\)
Paasche | `method = "Paasche"` | \({\Delta} C_{x_{1}} = \sum_{i} \left(x_{1i}^{T} - x_{1i}^{0}\right) x_{2i}^{T} ... x_{ni}^{T}\)
Marshall-Edgeworth | `method = "Marshall-Edgeworth"` | \({\Delta} C_{x_{1}} = \sum_{i} \left(x_{1i}^{T} - x_{1i}^{0}\right) \frac{x_{2i}^{T} + x_{2i}^{0}}{2} ... \frac{x_{ni}^{T} + x_{ni}^{0}}{2}\)
Walsh | `method = "Walsh"` | \({\Delta} C_{x_{1}} = \sum_{i} \left(x_{1i}^{T} - x_{1i}^{0}\right) \sqrt{x_{2i}^{T} x_{2i}^{0}} ... \sqrt{x_{ni}^{T} x_{ni}^{0}}\)

where 

\({\Delta} C_{x_{1}}\) is carbon emission change due to the change of factor \({x_{1}}\);

\(i\) is the subscript of a certain type of fuel or industrial sector; 

\(L(a,b)=\frac{a-b}{\ln{a} - \ln{b}}\) is a function defined by [Ang (2005)][2];

\(C_{i}^{0}\) and \(C_{i}^{T}\) are carbon emissions at a former time \(T\) (e.g. the first year) and a later time \(0\) (e.g. the second year), respectively; 

\(x_{1i}^{0}\) and \(x_{1i}^{T}\) are factor values of \(x_{1i}\) at a former time \(T\) (e.g. the first year) and a later time \(0\) (e.g. the second year), respectively.

**EDA** package includes a `CarbonEmission` dataset, which consists of:

- `carbon`: A data frame of carbon emissions from four types of fuels and three building sectors;

- `energy`: A data frame of energy consumption from four types of fuels and three building sectors;

- `buildingarea`: A data frame of annual building areas.

- `factordata`: A list of data frames of factors of carbon emission changes.

Further information can be found on the [manual][1] of **EDA** package.

## Converting variables to carbon emission factors

Generally, carbon emission factors are computed from a series of variables in terms of definitions of factors and research objectives. This section introduces the process of converting variables to carbon emission factors using the `CarbonEmission` dataset provided in the **EDA** package. If your factors data are available, please go to the next section. The result is the factors dataset `factordata`. If you want to directly try the `EDA` and `LMDI` functions, you can load the dataset `factordata`.

```{r}
library("EDA")
# load data
data("carbon")
data("energy")
data("buildingarea")
head(carbon)[1:3,]
```

```{r, eval = FALSE}
# calculate factors data
## Factor 1: Emission-factor effect (EF)
cdata <- carbon[,-c(1,2)]
edata <- energy[,-c(1,2)]
Sector <- c("b1", "b2", "b3")
xdata1 <- cdata/edata
## Factor 2: Energy-mix effect (EM)
Esum <- aggregate(edata, by=list(energy[,1]), FUN=sum, na.rm = TRUE)[,-1]
Esum <- Esum[rep(seq_len(nrow(Esum)), length(Sector)), ]
xdata2 <- edata/Esum
## Factor 3: Intensity effect (IE)
e_yearsum <- rowSums(Esum)
xdata3 <- Esum/e_yearsum
## Factor 4: Energy consumption density effect (DE)
area <- rep(buildingarea[,2], length(Sector))
xdata4 <- e_yearsum/area
xdata4 <- as.data.frame(xdata4)
## Factor 5: Area effect (AE)
xdata5 <- as.data.frame(area)
## list of factors
factordata <- list(xdata1, xdata2, xdata3, xdata4, xdata5)
names(factordata) <- c("EF","EM","IE","DE","AE")
```

## Log Mean Devisia Index (LMDI) method: `LMDI`

```{r}
data(carbon)
data(factordata)
## set parameters
cdata <- carbon[,-c(1,2)]
C0 <- cdata[1,]
CT <- cdata[2,]
X0 <- factordata[[2]][1,]
XT <- factordata[[2]][2,]
## run LMDI model
ed1 <- LMDI(C0, CT, X0, XT)
ed1
```

Further information can be found on the [manual][1] of **EDA** package.

## Energy consumption analysis: `EDA`

```{r}
data(carbon)
data(factordata)
## set parameters
cdata <- carbon[,-c(1,2)]
Year <- 2001:2005
Sector <- c("b1", "b2", "b3")
Fuel <- colnames(cdata)
Factor <- names(factordata)
## run EDA model
eda1 <- EDA(cdata, factordata, Year = Year, Factor = Factor, 
    Fuel = Fuel, Sector = Sector, method = "LMDI")
eda1
plot(eda1)
```

## Reference

Song Y (2018). EDA: Energy Decomposition Analysis. R package, https://CRAN.R-project.org/package=EDA.

Ang, B. W. (2005). The LMDI approach to decomposition analysis: a practical guide. Energy policy, 33(7), 867-871. doi: [10.1016/j.enpol.2003.10.010][2].

Liu, F. L., & Ang, B. W. (2003). Eight methods for decomposing the aggregate energy-intensity of industry. Applied Energy, 76(1-3), 15-23.

Marlay, R. C. (1984). Trends in industrial use of energy. Science, 226, 1277-1284.

Paasche, H. (1874). Uber die Preisentwicklung der letzten Jahre. Jahrbiicher fur Nationalokonomie und Statistik, 23, 168.

Marshall, A. (1887). Remedies for fluctuations of general prices.

Edgeworth, F. Y. (1925). Papers relating to political economy.

Walsh, C. M. (1921). The Problem of Estimation, a Seventeenth-century Controversy and Its Bearing on Modern Statistical Questions, Especially Index-numbers, by Correa Moylan Walsh.

## Author and Contact

Yongze Song

Email: yongze.song@postgrad.curtin.edu.au

Curtin University, Australia

[1]: https://CRAN.R-project.org/package=EDA
[2]: https://www.sciencedirect.com/science/article/pii/S0301421503003136
[3]: http://science.sciencemag.org/content/226/4680/1277
[4]: https://www.sciencedirect.com/science/article/pii/S0306261903000436#BIB2

